### 2025-10-06 thesis/design/initialization.tex
Improve the init section

\subsection{Initialization} % TODO: FIX THIS SECTION

Before the \textsc{Natural} system can process queries, it requires initialization
with historical data to construct a specific embedding space $v \in \mathcal{V}$ and
schema distance index $d \in \mathcal{D}$. This phase analyzes and samples existing
datasets and previous user interactions to build 
advanced example selection.

@@

Building upon DAIL-SQL's masked example selection, \textsc{Natural} utilizes 
cosine similarity in an embedding space $v$ to search for examples based on the
masked natural language query $q'$ and the respective masked SQL query $\omega'$.
This allows $\sigma$ to use cosine distance to search through historically observed
question answer pairs; eg, from previous user interactions or prevalent training
datasets like \textsc{Spider}, \textsc{Bird} or \textsc{SynSQL-2.5m}.

We therefore embedd a set of masked samples $\mathcal{S}$ into an embedding space $v_{\mathcal{S}}$,
where $\mathcal{S} = \{(q_1, s_1), ..., (q_i, s_i)\}$, $q$ is the natural language
question and $s$ is the corresponding SQL query.

The embedding space $v_{\mathcal{S}}$ can be defined as:

$$v_{\mathcal{S}} = \{~(q, \iota(mask(q)), s, \iota(mask(s)))~|~(q, s) \in \mathcal{S}~\}$$

Where the masking function $mask$ is defined as described in the sections~\ref{design:nl-masking}
and~\ref{design:sql-masking} and the embedding function is defined as $\iota$.
% TODO: Specify embedding model

@@

Masking of natural language is the process of replacing all words which are not
deemed \textit{structurally relevant} with the \texttt{<mask>} token. All words
$w$ in a constant whitelist $\mathcal{W}$ are determined structurally relevant.
% TODO: Fix typo
Estimating a constant $\mathcal{W}$ is described in section~\ref{implementation:whitelisting}.
% TODO: Fix reference to implementation section

@@

% TODO: Discuss potential limitations of masked embedding approach

% ALIGNMENT IMPROVEMENTS NEEDED:
% TODO: Add formal mathematical definition using enumerate style like indexing section
% TODO: Use consistent tuple notation (q,s,ω,d) matching definitions section  
% TODO: Add cross-references to Section~\ref{design:indexing} for schema distances
% TODO: Include similarity scoring methodology with weights w_q, w_s, w_sql
% TODO: Add subsection on "Distance Calculation" parallel to indexing section
% TODO: Use consistent variable naming (mathcal{V} vs V vs v)
% TODO: Add example showing multi-dimensional scoring like indexing graphs
% TODO: Include legend/explanation of masking tokens like indexing legend

@@

To determine structural similarity of database schemas systematically, we propose
using graph representation of database schemas as a data definition language (DDL)
independent representation of the database structure. Choosing a graph
representation allows us to leverage established methods for measuring graph similarity
such as Wasserstein-Weisfeiler-Lehman graph kernels \citep{WWL}. % TODO: Add proper WWL citation

Given a database schema $s$ we define the corresponding graph $G_s = (V, E, \ell, w)$ as:

@@

% TODO: Clarify relationship between edge weights and WWL computation
% TODO: Add example showing how label priority affects similarity scores

@@

% TODO: Fix inconsistent notation: use consistent variable names (G_s vs Natural-graphs)
% TODO: Improve transition between graph definition and distance measurement sections

>>>>

\subsection{Initialization}

Before the \textsc{Natural} system can process queries, it requires initialization
with historical data to construct a specific embedding space $v \in \mathcal{V}$ and
schema distance index $d \in \mathcal{D}$. This preprocessing phase analyzes existing
datasets and previous user interactions to enable semantically-aware example selection.

The initialization process consists of two main components: (1) embedding generation
for semantic similarity computation, and (2) schema indexing for structural similarity
measurement. These components work together to support the example selection function $\sigma$.

@@

Building upon DAIL-SQL's masked example selection approach, \textsc{Natural} constructs
an embedding space $v$ to enable semantic similarity search. The system uses cosine
similarity to identify relevant examples based on masked representations of both
natural language queries and SQL statements, allowing efficient retrieval from
large collections of question-answer pairs.

We therefore embed a set of training samples $\mathcal{T} = \{(q_1, \omega_1), ..., (q_n, \omega_n)\}$
into an embedding space $v$, where each $q_i$ is a natural language question and $\omega_i$
is the corresponding SQL query.

The embedding space $v$ can be formally defined as:

$$v = \{~(q, \iota(mask(q)), \omega, \iota(mask(\omega)))~|~(q, \omega) \in \mathcal{T}~\}$$

where $mask$ represents the masking function described in sections~\ref{design:nl-masking}
and~\ref{design:sql-masking}, and $\iota$ is the embedding function that maps text to vector representations.

@@

Masking of natural language is the process of replacing all words which are not
deemed \textit{structurally relevant} with the \texttt{<mask>} token. All words
$w$ in a constant whitelist $\mathcal{W}$ are determined structurally relevant.
The constant $\mathcal{W}$ contains common structural SQL keywords and is determined empirically.

@@

@@

To determine structural similarity of database schemas systematically, we propose
using graph representation of database schemas as a data definition language (DDL)
independent representation of the database structure. Choosing a graph
representation allows us to leverage established methods for measuring graph similarity
such as Wasserstein-Weisfeiler-Lehman graph kernels \citep{WWL}. 
Given a database schema $s$ we define the corresponding graph $G_s = (V, E, \ell, w)$ as:

@@

@@
